Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-07-05 main.tex
Changed simulation specs in text Reran simulations and obtained new figures (1000 runs for error plots, 10 runs for timing plots) Fixed couple of spelling things

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its formal security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

\section{Simulation and Results} \label{sec:sim_and_results}
As well as having shown the theoretical backing for the security of our scheme, we have simulated the proposed localisation method to evaluate its performance. A two-dimensional, linear, constant velocity process model,
\begin{equation*}
    \vec{x}_{k} = 
    \begin{bmatrix}
        1 & 0 & 0.5 & 0\\
        0 & 1 & 0 & 0.5\\
        0 & 0 & 1 & 0\\
        0 & 0 & 0 & 1
    \end{bmatrix} \cdot \vec{x}_{k-1} + \vec{w}_k\,,
\end{equation*}
where noise term $\vec{w}_k \sim \mathcal{N}(\vec{0}, \mat{Q})$ and
\begin{equation*}
    \mat{Q} = \frac{1}{10^3} \cdot
    \begin{bmatrix}
        0.4 & 0 & 1.3 & 0\\
        0 & 0.4 & 0 & 1.3\\
        1.3 & 0 & 5.0 & 0\\
        0 & 1.3 & 0 & 5.0
    \end{bmatrix}\,,
\end{equation*}
was simulated and tracked with the algorithms in section \ref{subsec:pseudocode}, using a linear Kalman filter for the navigator's local state prediction. Code was written in the C programming language, using the MPI library \cite{theopenmpiprojectOpenMPI2020} to support asynchronous computations by the sensors and navigator. The MG1 mask generation function and the SHA256 hash function, from the OpenSSL library \cite{theopensslprojectOpenSSL2020}, were used to implement the required hash function $H$, and the Libpaillier library \cite{bethencourtLibpaillier2010} was used for the Paillier encryption scheme. Additionally, GNU libraries, GSL \cite{thegsldevelopmentteamGSLGNUScientific2019} and GMP \cite{granlundGMPGNUMultiple2020}, were used for algebraic operations and multiple-precision encoded integers, respectively. All execution was performed on a 3.00GHz Intel i7-9700 CPU, running on the Windows Subsystem for Linux (WSL).

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. These layouts of varying sensor distances are shown next to the simulation initial state and a sample track in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Different simulation layouts with varying distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, that is algebraically equivalent to the EKF, typically used in industry for linearising non-linear state estimation. Measurement and estimation in each layout from figure \ref{fig:sim_layouts} consists of $50$ filter iterations and was run $100$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{Average RMSE of our privacy-preserving filter and the standard EIF for different layouts.}
    \vspace{-\baselineskip}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see the similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We can attribute the similarity in RMSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

In addition to filter error, computational performance is important to consider when relying on cryptographic methods. Figure \ref{fig:sim_timing} shows the average execution time of $50$ simulation runs with varying numbers of sensors, for different key sizes (bit lengths of $N$). Here, increasing the number of sensors primarily affects the number of inter-process communications and aggregation modular multiplications due to the asynchronous implementation.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/timing.pdf}
    \caption{Runtimes for varying key sizes and numbers of sensors.}
    \label{fig:sim_timing}
\end{figure}
We can see from the figure that the predominant computational costs stem from cryptographic computations and are directly dependent on the chosen key size. The key size should be chosen such that sufficient security is achieved, and the current recommendation, when relying on the DCRA for security (difficulty of factorising $N$), is the use of $2048$ bit length keys \cite{barkerRecommendationPairwiseKey2019}. For our implementation of the filter, this results in a filter update computation duration of roughly $1.6s$.

>>>>

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and present an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its formal security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

\section{Simulation and Results} \label{sec:sim_and_results}
As well as having shown the theoretical backing for the security of our scheme, we have simulated the proposed localisation method to evaluate its performance. A two-dimensional, linear, constant velocity process model,
\begin{equation*}
    \vec{x}_{k} = 
    \begin{bmatrix}
        1 & 0 & 0.5 & 0\\
        0 & 1 & 0 & 0.5\\
        0 & 0 & 1 & 0\\
        0 & 0 & 0 & 1
    \end{bmatrix} \cdot \vec{x}_{k-1} + \vec{w}_k\,,
\end{equation*}
where noise term $\vec{w}_k \sim \mathcal{N}(\vec{0}, \mat{Q})$ and
\begin{equation*}
    \mat{Q} = \frac{1}{10^3} \cdot
    \begin{bmatrix}
        0.4 & 0 & 1.3 & 0\\
        0 & 0.4 & 0 & 1.3\\
        1.3 & 0 & 5.0 & 0\\
        0 & 1.3 & 0 & 5.0
    \end{bmatrix}\,,
\end{equation*}
was simulated and tracked with the algorithms in section \ref{subsec:pseudocode}, using a linear Kalman filter for the navigator's local state prediction. Code was written in the C programming language using the MPI library \cite{theopenmpiprojectOpenMPI2020} to support asynchronous computations by the sensors and navigator. The MG1 mask generation function and the SHA256 hash function, from the OpenSSL library \cite{theopensslprojectOpenSSL2020}, were used to implement the required hash function $H$, and the Libpaillier library \cite{bethencourtLibpaillier2010} was used for the Paillier encryption scheme. Additionally, GNU libraries, GSL \cite{thegsldevelopmentteamGSLGNUScientific2019} and GMP \cite{granlundGMPGNUMultiple2020}, were used for algebraic operations and multiple-precision encoded integers, respectively. All execution was performed on a 3.33GHz Xeon W3680 CPU, running on the Windows Subsystem for Linux (WSL).

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. These layouts of varying sensor distances are shown next to the simulation initial state and a sample track in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Different simulation layouts with varying distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, which is algebraically equivalent to the EKF typically used in industry for linearising non-linear state estimation. Estimation in each layout from figure \ref{fig:sim_layouts} consisted of $50$ filter iterations and was run $1000$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{Average RMSE of our privacy-preserving filter and the standard EIF for different layouts.}
    \vspace{-\baselineskip}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see a strong similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We attribute this similarity in RMSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

In addition to filter error, computational performance is important to consider when relying on cryptographic methods. Figure \ref{fig:sim_timing} shows the averages of $10$ execution times when varying the numbers of sensors and key sizes (bit lengths of $N$). Here, increasing the number of sensors primarily affected the number of inter-process communications and aggregation modular multiplications due to the asynchronous implementation.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/timing.pdf}
    \caption{Runtimes for varying key sizes and numbers of sensors.}
    \label{fig:sim_timing}
\end{figure}
We can see from the figure that the predominant computational costs stem from cryptographic computations and are directly dependent on the chosen key size. The key size should be chosen such that sufficient security is achieved, and the current recommendation, when relying on the DCRA for security (difficulty of factorising $N$), is the use of $2048$ bit length keys \cite{barkerRecommendationPairwiseKey2019}. For our implementation of the filter, run on the aforementioned hardware, this results in a filter update computation duration of roughly $1.7s$.
